---
title: "Reading csv files with accelerometer raw data with GGIR"
output: rmarkdown::html_vignette
    toc : true
    number_sections: true
    toc_depth: 3
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{Reading csv files with accelerometer raw data with GGIR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

If you want to use GGIR but the accelerometer brand which you collected data with is not supported by GGIR, the `read.myacc.csv` function may be of help. This function is able to read data from a csv containing the raw data from tri-axial accelerometers. The term _**raw**_ refers to data being expressed in m/s<sup>2</sup> or gravitational acceleration. Thus, the usability of GGIR is extended to almost any device as long as a single csv file with the raw acceleroemter data for each of the three axes is provided.

**How it works:**

Internally GGIR loads csv files with accelerometer data and standardises the output format to make the data compatible with other GGIR functions. Data format standardization includes unit of measurement (_G_).

```{r setup}
library(GGIR)
```
